Data exploration, quality control and statistical analysis of ChIP-exo experiments

### 2016-03-17 rmarkdown/paper.Rnw
updated GC content + mappability arguments

We evaluated ChIP-exo data for CTCF factor from human genome
\cite{exo1} to investigate issues specific to eukaryotic genomes for
binding sites identification. Figures \ref{fig:comp}C and
\ref{fig:comp}D display the bin-level average read counts against
mappability and GC content. Each data point is obtained by averaging
the read counts across bins with the same mappability of GC
content. In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
counts linearly increases with the mappability score and in Figure
\ref{fig:comp}D it is shown that for GC - content below 0.6, the mean
ChIP tag count increases and for GC - content greater than 0.6 it
shows a decreasing trend.

These results indicate that binding site identification in ChIP-exo
sample might also benefit from the use of methods that take into
account of apparent sequence biases such as mappability and GC
content.

@@

where $\delta_i$ is the indicator if the base at coordinate $i$ can be
mapped uniquely by a 32 bp sequence at position $i$, and $L$ is the
expected fragment length. GC - content score is defined analogously,
where $\delta_i$ represents the ocurrence of a G or C at the $i$-th
position in the genome.

>>>>

We evaluated ChIP-exo data for CTCF factor from human genome
\cite{exo1} to investigate issues specific to eukaryotic genomes for
binding sites identification. Figures \ref{fig:comp}C and
\ref{fig:comp}D display the bin-level average read counts against
mappability and GC content. Each data point is obtained by averaging
the read counts across bins with the same mappability of GC -
content. In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
counts linearly increases with the mappability score and in Figure
\ref{fig:comp}D it is shown that for GC - content below 0.6, the mean
ChIP tag count increases and for GC - content greater than 0.6 it
shows a decreasing trend. Kuan et al., 2011 \cite{mosaics} studied the
presence of the mappability and GC - content biases in ChIP-Seq's
background. It is not surprising to see these biases also present in
ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to be linearly
correlated for enriched regions (Figure \ref{fig:comp}A). Rozowsky et
al., 2009 \cite{peakseq} and Valouev et al., 2008 \cite{quest} provide
in depth analysis of the mappability and GC - content biases for
ChIP-Seq respectively. Finally, these results indicate that binding
site identification for ChIP-exo data benefits from using methods that
take into account of apparent sequence biases such as mappability and
GC content.

@@

where $\delta_i$ is the indicator if the base at coordinate $i$ can be
mapped uniquely by a 32 bp sequence at position $i$, and $L$ is the
expected fragment length. GC - content score is defined analogously,
where $\delta_i$ represents the occurrence of a G or C at the $i$-th
position in the genome.
